---
title: "Example analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pheatmap)
library(here)
library(tidyverse)
library(stringr)
library(lubridate)
library(tidytuesdayR)
```


## Part 1C: Create an example data analysis

### 3C.1.

**Question**: How do coffee ratings vary by the species of coffee bean (arabica or robusta) and location?

**Original data** The data are from the Coffee Quality Database courtesy of Buzzfeed Data Scientist James LeDoux. <https://github.com/jldbc/coffee-quality-database>

**Data directory**: The data directories are accessible via TidyTuesday. <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-07/readme.md>


### 3C.2

```{r}
# tests if a directory named "data" exists locally
if(!dir.exists(here("data"))) { dir.create(here("data")) }

# saves data only once (not each time you knit a R Markdown)
if(!file.exists(here("data", "coffee_ratings.csv"))) {
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
write_csv(coffee_ratings, file = here("data", "coffee_ratings.csv"))
}

# read in the data locally each time you knit/render
coffee_ratings <- read_csv(here("data", "coffee_ratings.csv"))
```

### 1C.3. Data wrangling and visualization

```{r}
coffee_ratings$date <- coffee_ratings$grading_date %>%
  mdy()

coffee_ratings %>%
  ggplot(aes(date, total_cup_points)) +
  geom_smooth(aes(color = species)) +
  labs(
    title = "Coffee bean ratings over time",
    subtitle = "Arabica and robusta beans", 
    caption = "2010 - 2018",
    x = "Time (Years)", 
    y = "Cup points") +
  labs(color = "Bean species") +
  theme_minimal() +
  theme(legend.background=element_blank(), 
        legend.key = element_blank())

```


```{r}
coffee_ratings$year <- coffee_ratings$date %>%
  year()

coffee_ratings %>%
  filter(year >= 2012, year <= 2017) %>%
  group_by(country_of_origin) %>%
  summarise(total_points = sum(total_cup_points))%>%
  drop_na() %>%
  arrange(desc(total_points)) %>%
  slice(1:25) %>%
  mutate(country_order = fct_reorder(country_of_origin, total_points)) %>%
  ggplot(aes(country_order, total_points)) +
  geom_col() +
  coord_flip()+
  labs(
    title = "Top 25 coffee bean countries",
    subtitle = "2012 - 2017",
    caption = "Arabica and robusta beans combined",
    y = "Cup points",
    x = "Country") +
  theme_minimal()
```

```{r}
coffee_ratings %>%
  filter(year >= 2012, year <= 2017) %>%
  filter(country_of_origin == "Mexico" | country_of_origin == "Guatemala" | country_of_origin =="Colombia" |
         country_of_origin == "Brazil"| country_of_origin == "Taiwan") %>%
  group_by(country_of_origin, species) %>%
  summarise(total_points = sum(total_cup_points)) %>%
  arrange(desc(total_points))
  
```

```{r}
coffee_species <- coffee_ratings %>%
  split(coffee_ratings$species)

coffee_species %>%
  map_dbl(.f = ~mean(.x$total_cup_points))

coffee_ratings %>%
  select(c("aroma", "flavor", "aftertaste", "acidity", "body", "balance", "uniformity", "sweetness"))%>%
  split(coffee_ratings$species) %>%
  map(.f = ~colMeans(.x))
```


```{r}
coffee_ratings_long <- coffee_ratings %>%
  pivot_longer(
    cols = c("aroma", "flavor", "aftertaste", "acidity", "body", "balance", "uniformity", "sweetness"),
    names_to = "rating_name", 
    values_to = "rating_grade")

coffee_ratings_long %>%
  filter(year >= 2012, year <= 2017, rating_grade > 2) %>%
  group_by(species) %>%
  ggplot(aes(species, rating_grade)) +
  geom_violin(aes(fill = species)) +
  facet_wrap(vars(rating_name)) +
  labs(
    title = "Coffee attributes by bean species",
    subtitle = "Arabica and robusta beans", 
    caption = "2012 - 2017",
    x = "Bean species", 
    y = "Rating grade (points)") +
  labs(fill = "Bean species") +
  theme_minimal() +
  theme(legend.background=element_blank(), 
        legend.key = element_blank())

```

```{r}
coffee_heatmap <- coffee_ratings %>%
  filter(species == "Arabica") %>%
  select(c("country_of_origin", "aroma", "flavor", "aftertaste", "acidity", "body", "balance", "uniformity", "sweetness")) %>%
  group_by(country_of_origin) %>%
  summarise(aroma_rating = mean(aroma),
            flavor_rating = mean(flavor),
            aftertaste_rating = mean(aftertaste), 
            acidity_rating = mean(acidity), 
            body_rating = mean(body),
            balance_rating = mean(balance), 
            uniformity_rating = mean(uniformity), 
            sweetness_rating = mean(sweetness)) %>%
  drop_na()
  

names(coffee_heatmap) <- c("country_of_origin", "aroma", "flavor", "aftertaste", "acidity", "body", "balance", "uniformity", "sweetness")

coffee_heatmap <- column_to_rownames(coffee_heatmap, var = "country_of_origin")

pheatmap(coffee_heatmap)

pheatmap(coffee_heatmap, cluster_rows = F, cluster_cols = F)

pheatmap(coffee_heatmap, cluster_rows = T, cluster_cols = F)


```

### 1C.5 Results summary
Arabica coffee beans appear to be more popular and have higher ratings overtime. Additionally, the five countries with the highest coffee ratings only produce arabica beans. 

### 1C.6. Functions used

#### dplyr
(1) group_by
(2) summarise
(3) arrange
(4) slice
(5) filter
(6) mutate

#### tidyr
(1) pivot_longer
(2) drop_na

#### forcats
(1) fct_reorder

#### lubridate
(1) mdy
(2) year

#### purrr
(1) map
(2) map_dbl

#### ggplot2
(1) geom_smooth
(2) geom_violin
(3) geom_col

#### pheatmap
(1) pheatmap
(2) cluster_rows
(3) cluster_cols
